---
title: "Assignment 3"
author: "gwilly4280"
date: "26/01/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Project Info

Written by: **gwilly4280**

Date Uploaded: **29-01-2021**

Link to repository: **https://github.com/gwilly4280/BIOL432** 

### Loading Libraries & Data

For this we'll be using *dplyr* for easy data-wrangling
```{r, message = F, warning = F}
library(dplyr)
```

Loading & quickly displaying summary stats for the data
```{r}
myData <- read.csv("./InputData/fallopiaData.csv")
summary(myData)
```

### Data wrangling

**Code which performs the following:**

* Remove rows with ‘Total’ biomass < 60

* Reorder the columns so that they are in the order: ‘Total’, ‘Taxon’, ‘Senario’, ‘Nutrients’, and remove the other columns

* Make a new column TotalG, which converts the ‘Total’ column from mg to grams AND replace Total with TotalG, and add it to the dataset.

```{r}
# Replace dataframe w/ cleaned data
myData <- myData %>% 
   # Remove data with biomass < 60
   filter(Total >= 60) %>% 
   # Reorder & keep only Total, Taxon, Scenario, Nutrients
   transmute(TotalG = Total/1000, # Replace Total w/ TotalG
             Taxon,
             Scenario,
             Nutrients)

# Displaying head of resulting data-table
head(myData)
```


### Custom Function

This function takes a numeric vector and a String as input, and outputs a single numeric value. Output is dependent on String identity: if the supplied String is *"Average"*, will calculate and return the **mean** of the supplied vector; if String is *"Sum"*, will calculate and return the **sum** of the values within the supplied vector; if the String is *"Observations"*, will return the **length** of the supplied vector. If the supplied string **does not match** any of the above cases, function will **raise a custom exception**.
```{r}
process_vector <- function(vect, string){
   # Chained if-statements based on String Identity
   if (string == "Average"){
      output <- mean(vect)
   } else if (string == "Sum"){
      output <- sum(vect)
   } else if (string == "Observations"){
      output <- length(vect)
   } else {
      # Raise exception if String invalid
      stop("INVALID INPUT: String must be one of: 'Average', 'Sum', or 'Observations'")
   }
   # Returning output calculated from within the chained ifs
   return(output)
}
```

### Error checking:

Checking Average
```{r}
# Setting Vector
vect <- c(1:10)

# Checking Average (should output 5.5)
process_vector(vect, "Average")
```

Checking Sum
```{r}
# Checking Sum (should output 55)
process_vector(vect, "Sum")
```

Checking Observations
```{r}
# Checking Observations (should output 10)
process_vector(vect, "Observations")
```

Checking purposeful incorrect String (custom exception will be raised)
```{r, error=TRUE}
# Error checking incorrect string
process_vector(vect, "Not the right string")
```

Checking type error in String Input (custom exception will be raised)
```{r, error = TRUE}
# Error checking a type-error
process_vector(vect, 7)
```

### Using Custom Function

Write some R code that uses your function to count the total number of observations in the 'Taxon' column.

```{r}
process_vector(myData$Taxon, "Observations")
```

Write some R code that uses your function  to calculate the average TotalG for each of the two Nutrient concentrations.
```{r, message = F}
# Storing table to variable
sumTable <- myData %>% 
   # Grouping by Nutrient Levels
   group_by(Nutrients) %>% 
   # Producing Summary Table using custom function
   summarize(MeanTotalG = process_vector(TotalG, "Average"))

# Using knitr function for cleaner output
knitr::kable(sumTable, align = "cc")
```


### Write to File

Write the new data to a file called "WrangledData.csv" in the Output folder. 
```{r}
# Write to file
write.csv(myData, 
          file = "./Output/WrangledData.csv", # Set path
          row.names = FALSE) # Not saving row numbers to csv for cleaner data

# Check if file-write was successful
head(read.csv("./Output/WrangledData.csv"))
```



